%\chapter*{Acknowledgments}
\centerline{{\bf \Large Acknowledgments}} \vspace{5mm} \noindent

I would like to express my foremost gratitude to my supervisor, Prof. Fugee Tsung,
for his insightful guidance and persistent support. Without him, it couldn't be
possible that I have this chance to study here and pursue my Ph.D. degree. I have
benefited a lot from his character as a father in academia. He pointed out a
fundamental research topic, which made this thesis possible. Besides advising my
research, he provided me many opportunities in participating academic conferences,
writing research proposals, and involving industrial projects.

I am grateful to my thesis committee members, including Prof. Richard H. Y. So and
Prof. Emily Au in the IELM department, Prof. Xinghua Zheng in the ISOM department,
Prof. Kam Sing Wong in the PHYS department, and Prof. Min Xie in Department of
Systems Engineering and Engineering Management (SEEM) at City University of Hong
Kong. I also thank very much Prof. Ning Cai and Prof. Jiheng Zhang for serving on
the committee of my prelim dissertation. In addition, I express my sincere thanks to
Prof. Jye-Chyi Lu from School of Industrial and Systems Engineering at Georgia
Institute of Technology and Ms. Fona Wong in the IELM department, who provided great
help for my defense.

There are special thanks to Prof. Changliang Zou from Nankai University, who is an
excellent academic elder brother. He provided lots of valuable suggestions on my
research. I enjoy discussing research problems and details with him. I also express
my gratitude to Prof. Wei Jiang from Shanghai Jiaotong University, who provided many
guidelines in my academic career.

Besides, I would like to thank Prof. Rachel Zhang and Prof. Qian Liu in the IELM
department, Prof. Bingyi Jing and Prof. Shiqing Ling in the MATH department, as well
as Prof. Mike So in the ISOM department for the courses they offered, which impacted
and motivated my research. I would like to thank all my friends and fellow graduate
students in HKUST, especially those in Quality Lab.

I am forever indebted to my father and mother for their love. There are no words
that can express my appreciation for their greatest care and favor. I am always
thinking of them who are at home a thousand miles away.
