\addcontentsline{toc}{chapter}{Bibliography}


\begin{thebibliography}{}

\item Agresti, A. (2002) {\it Categorical Data Analysis}, 2nd
edition, Wiley, New York.

\item Agresti, A. (2010) {\it Analysis of Ordinal Categorical Data},
2nd edition, Wiley, New York.

\item Anderson, T.W. (2003) {\it An Introduction to Multivariate
Statistical Analysis}, 3rd edition, Wiley, New York.

\item Ben-Gal, I., Morag, G. and Shmilovici, A. (2003) Context-based statistical
process control: a monitoring procedure for state-dependent processes. {\it
Technometrics}, 45, 293--311.

\item Bersimis, S., Psarakis, S. and Panaretos, J. (2007) Multivariate
statistical process control charts: an overview. {\it Quality and Reliability
Engineering International}, 23, 517--543.

\item Bishop, Y.M.M., Fienberg, S.E. and Holland, P.W. (2007) {\it Discrete Multivariate
Analysis}, Springer, New York.

\item Brice, P. and  Jiang, W. (2009) A context tree  method  for multistage
fault detection and isolation with applications to commercial video broadcasting
systems. {\it IIE  Transactions}, 41, 776--789.

\item Brice, P., Jiang, W. and Wan, G.H. (2012) A cluster-based context-tree
model for multivariate data streams with applications to anomaly detection. {\it
INFORMS Journal on Computing}, forthcoming.

\item Chiu, J. and Kuo, T. (2008) Attribute control chart for multivariate Poisson
distribution. {\it Communications in Statistics: Theory and Methods}, 37, 146--158.

\item Christensen, R. (1997) {\it Log-Linear Models and Logistic Regression},
2nd edition, Springer, New York.

\item Cs\"{o}rg\H{o}, M. and Horv\'{a}th, L. (1997) {\it Limit Theorems in Change-Point
Analysis}, John Wiley \& Sons, New York.

\item Dahinden, C., Parmigiani, G., Emerick, M.C. and B\"uhlmann, P.
(2007) Penalized likelihood for sparse contingency tables with an application to
full-length cDNA libraries. {\it BMC Bioinformatics}, 8, 476.

\item Fienberg, S.E. and Rinaldo, A. (2007) Three centuries of categorical data analysis:
log-linear models and maximum likelihood estimation. {\it Journal of Statistical
Planning and Inference}, 137, 3430--3445.

\item Garvin, D.A. (1987) Competing in the eight dimensions of quality. {\it Harvard Business
Review}, 87, 101--109.

\item Hawkins, D.M. and Qiu, P. (2003) The changepoint model for statistical process
control. {\it Journal of Quality Technology}, 35, 355--366.

%\item Jensen, W.A., Jones, L.A., Champ, C.W. and Woodall, W.H. (2006) Effects of parameter estimation on control chart
%properties: a literature review. {\it Journal of Quality
%Technology}, 38, 349--364.

\item Johnson, N.L., Kotz, S. and Balakrishnan, N. (1997) {\it Discrete Multivariate
Distributions}, Wiley, New York.

\item Kaminsky, F.C., Benneyan, J.C., Davis, R.D. and Burke, R.J.
(1992) Statistical control charts based on a geometric distribution. {\it Journal of
Quality Technology}, 24, 63--69.

\item Lowry, C.A. and Montgomery, D.C. (1995) A review of multivariate control charts.
{\it IIE Transactions}, 27, 800--810.

\item Lowry, C.A., Woodall, W.H., Champ, C.W. and Rigdon, S.E. (1992) A multivariate exponentially
weighted moving average control chart. {\it Technometrics}, 34, 46--53.

\item Lu, X.S., Xie, M., Goh, T.N. and Lai, C.D. (1998) Control charts
for multivariate attribute processes. {\it International Journal of Production
Research}, 36, 3477--3489.

\item Lucas, J.M. and Saccucci, M.S. (1990) Exponentially
weighted moving average control schemes: properties and enhancements. {\it
Technometrics}, 32, 1--29.

\item Marcucci, M. (1985) Monitoring multinomial
processes. {\it Journal of Quality Technology}, 17, 86--91.

\item McCool, J.I. and Joyner-Motley, T. (1998) Control charts applicable when the
fraction nonconforming is small. {\it Journal of Quality Technology}, 30, 240--247.

\item McCullagh, P. and Nelder, J.A. (1989) {\it Generalized Linear Models}, 2nd
edition, Chapman \& Hall/CRC, London.

\item Montgomery, D. C. (2009) {\it Statistical Quality Control: A Modern Introduction}, 6th
edition, Wiley, New York.

\item Nelson, L.S. (1994) A control chart for parts-per-million nonconforming
items. {\it Journal of Quality Technology}, 26, 239--240.

\item Niaki, S.T.A. and Abbasi, B. (2007) On the monitoring of
multi-attribute high-quality production processes. {\it Metrika}, 66, 373--388.

\item Ning, X. and Tsung, F. (2012) Improved design of kernel-distance-based charts
using support vector methods. {\it IIE Transactions}, accepted.

\item Patel, H.I. (1973) Quality control methods for
multivariate binomial and Poisson distributions. {\it Technometrics}, 15, 103--112.

\item Ryan, A.G., Wells, L.J. and Woodall, W.H. (2011) Methods for monitoring
multiple proportions when inspecting continuously. {\it Journal of Quality
Technology}, 43, 237--248.

%\item Serfling, R.J. (1980) {\it Approximation Theorems of
%Mathematical Statistics}, John Wiley \& Sons, New York.

\item Simes, R.J. (1986) An improved Bonferroni procedure for multiple tests of
significance, {\it Biometrika}, 73, 751--754.

\item Srivastava, M.S. and Worsley, K.J. (1986) Likelihood ratio tests for a change in
the multivariate normal mean, {\it Journal of the American Statistical Association},
81, 199--204.

\item Sun, R. and Tsung, F. (2003) A kernel-distance-based multivariate control chart
using support vector methods. {\it International Journal of Production Research},
41, 2975--2989.

\item Tibshirani, R. (1996) Regression shrinkage and selection via the lasso. {\it
Journal of the Royal Statistical Society, Series B}, 58, 267--288.

\item Topalidou, E. and Psarakis, S. (2009) Review of multinomial and
multiattribute quality control charts. {\it Quality and Reliability Engineering
International}, 25, 773--804.

\item Wang, K. and Jiang, W. (2009) High-dimensional process monitoring and fault
isolation via variable selection. {\it Journal of Quality Technology}, 41, 247--258.

\item Wang, K., Jiang, W., and Tsung, F. (2012) A variable-selection-based multivariate
EWMA chart for process monitoring and diagnosis. {\it Journal of Quality
Technology}, tentatively accepted.

\item Woodall, W.H. and Ncube, M.M. (1985) Multivariate CUSUM quality control
procedures. {\it Technometrics}, 27, 285--292.

\item Woodall, W.H. (1997) Control charts based on attribute data: bibliography and
review. {\it Journal of Quality Technology}, 29, 172--183.

\item Wu, C. F. J. and Hamada, M. (2000) {\it Experiments: Planning, Analysis, and
Parameter Design Optimization}, Wiley, New York.

\item Yang, Z., Xie, M., Kuralmani, V. and Tsui, K.L. (2002) On the performance
of geometric charts with estimated control limits. {\it Journal of Quality
Technology}, 34, 448--458.

\item Yao, Y. (1988) Estimating the number of change-points via Schwarz criterion.
{\it Statistics and Probability Letters}, 6, 181--189.

%\item Zou, C., Tsung, F. and Wang, Z. (2007) Monitoring general linear
%profiles using multivariate EWMA schemes. {\it Technometrics}, 49, 395--408.

\item Zou, C., Jiang, W. and Tsung, F. (2011) A LASSO-based SPC diagnostic framework for
multivariate statistical process control. {\it Technometrics}, 53, 297--309.

\item Zou, C. and Qiu, P. (2009) Multivariate statistical process control using
LASSO. {\it Journal of the American Statistical Association}, 104, 1586--1596.

\item Zou, C. and Tsung, F. (2008) Directional MEWMA schemes for
multistage process monitoring and diagnosis. {\it Journal of Quality Technology},
40, 407--427.

\item Zou, C. and Tsung, F. (2011) Multivariate sign EWMA control
chart. {\it Technometrics}, 53, 84--97.

\item Zou, C., Tsung, F. and Liu, Y. (2008) A change point approach for Phase I
analysis in multistage processes. {\it Technometrics}, 50, 344--356.

\item Zou, C., Wang, Z., and Tsung, F. (2012) A spatial rank-based multivariate EWMA
control chart. {\it Naval Research Logistics}, 59, 91--110.

\end{thebibliography}
